---
title: "p8105_hw3_as6445"
author: "Ayako Sekiya"
date: "2022-10-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
data("instacart")
data("ny_noaa")
```

## Problem 1

There are `r nrow(instacart)` observations and `r ncol(instacart)` columns in the Accel data. The variables included are as following: `r colnames(instacart)`.  


## Problem 2

```{r tidy accel}
accel = 
  read_csv(file = "./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity",
    values_to = "activity_count") %>% 
  mutate(day_type= ifelse(day=='Saturday','Weekend',
                   ifelse(day=='Sunday','Weekend','Weekday')))
      
```
There are `r nrow(accel)` observations and `r ncol(accel)` columns in the Accel data. The variables included are as following: `r colnames(accel)`.  
```{r analysis}
accel_analysis = accel %>% 
  group_by(day_type) %>% 
  summarize(sum_totalactivity = sum(activity_count))
```

### Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r aggregate}
accel_24 = accel %>% 
  select (day, activity, activity_count, week) %>% 
  
  ggplot(aes(x =activity , y = activity_count, color=day)) + 
  geom_point()+
   geom_line(alpha = 0.5 ) +  
  labs(
    title = "24-hour activity time for each day by week",
    x = "day",
    y = "activity count (min))",
    caption = "Data from accel data") +
  theme(plot.title = element_text(size = 10, face = "bold"),
        legend.position = "right",
        legend.text = element_text(size = 7))+
  scale_x_discrete(breaks = c(1980, 1985, 1990, 1995, 2000, 2005, 2010)) +
  theme(plot.title = element_text(size = 10, face = "bold"),
        legend.position = "right",
        legend.text = element_text(size = 7))  

ggsave("accel_24.pdf")
```

## Problem 3

We will first tidy the dataset!

```{r tidy noaa}
ny_noaa <- ny_noaa %>% 
  janitor::clean_names() %>%
  separate(date, into=c("year", "month", "day"), sep= "-") %>% 
  mutate(tmax = as.numeric(tmax, na.rm = TRUE)/10,
         tmin = as.numeric(tmin, na.rm = TRUE)/10,
        prcp = as.numeric(prcp, na.rm = TRUE)/10) %>% 
  mutate(month= as.character(month.name[month]))
```
   
For snowfall, what are the most commonly observed values? Why?

To answer this question, we will use `group_by` and `summarize`. 

```{r snowfall}
snow <- ny_noaa %>% 
  group_by(snow) %>% 
  summarize(n_obs = n())
```

The most commonly observed values is 0 based on this output which would make sense because there should be more days that it does not snow in NY than days that snows. 

### Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r two-panel avg max temp}
avg_jan_july_p = 
  ny_noaa %>% 
  select(id, month, year, tmax, tmin) %>% 
  filter(month=='January'| month=='July') %>%
  group_by(id, month, year) %>% 
  summarize(mean_tmax = mean(tmax), ) %>%
  ggplot(aes(x = year, y = mean_tmax, color = id)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none") + 
  labs(
    title = "Average max temps in January and July",
    x = "Years",
    y = "Average Max daily temperature (C)",
    caption = "Data from the ny noaa package") +
  scale_x_continuous(breaks = c(1980, 1985, 1990, 1995, 2000, 2005, 2010)) +

ggsave("avg_jan_july_p.pdf")
```

### Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option). 

```{r two-panel tmin and max}
tmax_tmin_p = 
  ny_noaa %>% 
  select(tmax, tmin, year) %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex() + 
  labs(
    title = "Temperature max/ minplot",
    x = "Maximum daily temperature (C)",
    y = "Minimum daily temperature (C)",
    caption = "Data from the ny noaa package")+
  scale_x_continuous(
    breaks = c(-30, -15, 0, 15, 30, 45, 60), 
    labels = c("-30°C","-15°C", "0°C", "15°C", "30°C", "45°C", "60°C")) +
  scale_y_continuous(
    breaks = c(-60, -45, -30, -15, 0, 15, 30, 45, 60), 
    labels = c("-60°C","-45°C","-30°C","-15°C", "0°C", "15°C", "30°C", "45°C", "60°C")) 

ggsave("tmax_tmin_p.pdf")
```

From this plot, we can see that there are the highest concentrations where there is the lightest color. There seems to be a correlation between minimum and maximum daily temperatures around where there is the most concentrated data. 

### Make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
snow_plot<- ny_noaa %>% 
  select (year, snow) %>% 
  filter(snow > 0 & snow < 100) %>% 
  mutate(year = as.character(year)) %>% 
  ggplot(aes(x = year, y = snow, fill = year)) + 
  geom_violin(alpha = 0.5 ) +  
  labs(
    title = "Snowfall distribution (0-100mm) by year",
    x = "Year",
    y = "Snowfall (mm)",
    caption = "Data from the ny noaa package") +
  scale_x_discrete(breaks = c(1980, 1985, 1990, 1995, 2000, 2005, 2010)) +
  theme(plot.title = element_text(size = 10, face = "bold"),
        legend.position = "right",
        legend.text = element_text(size = 7))

ggsave("snow_plot.pdf")
```

From this plot, there seems to be the highest concentration of snowfall between approximately 0-25mm. The data is skewed to the right, and there are fewer occasions where it snowed more than 75mm of snow throughout the years. 
